---
title: "Criminology Data Analysis Course: Semester Plan"
subtitle: "16-Week Learning Path"
author: "Dr. Kuralarasan Kumar"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    theme: flatly
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(knitr)
librar- **Student Feedback Loops**: 1-minute pulse polls at Weeks 1, 5, 8, 15(DT)
```

> **Note:** This is a rough plan and subject to change based on student needs, technical requirements, and institutional feedback throughout the semester.

# Course Overview

## Learning Objectives
By the end of this course, students will be able to:
- Navigate R/RStudio and manage reproducible projects
- Import, clean, and manipulate criminological datasets  
- Conduct descriptive and inferential statistical analyses
- Create publication-quality visualizations and APA-formatted tables
- Interpret results within criminological contexts
- Transition from SPSS workflows to R-based analysis

## Assessment Structure
```{r assessment, echo=FALSE}
assessment <- data.frame(
  Component = c("Onboarding + Quizzes", "Assignment 1 (Crime Rates)", 
                "Midterm Mini-Project", "Final Project", "Participation"),
  Weight = c("10%", "15%", "25%", "40%", "10%"),
  Description = c("Low-stakes skill checks", "APA table + report", 
                  "District analysis + peer review", "Student-chosen question",
                  "Clinics, feedback, reflection")
)
kable(assessment, caption = "Assessment Weights")
```

## Platform Integration
- **Dodona**: Auto-graded exercises with instant feedback
- **swirl**: Interactive R console tutorials  
- **Ufora**: Quizzes, resources, discussions, announcements
- **GitHub**: Code repositories and dataset access (read-only for students)

---

# Weekly Schedule

## Week 1: Onboarding & Baseline
**Topics:** R/RStudio setup, projects, Dodona submission, Ufora navigation, GitHub access

**Learning Activities:**
- Install R, RStudio, and create first project
- Complete Dodona onboarding item (basic syntax, variables)
- Short Ufora quiz on measurement levels in criminology
- Install swirl package and explore interface

**Deliverables:**
- Onboarding quiz completion
- Baseline self-efficacy poll (1-minute survey)

**Tracking Metrics:**
- % onboarding completed by Day 5
- Common error patterns → targeted hints added
- Initial confidence levels in statistics/programming

---

## Week 2: Data Structures & Measurement Levels
**Topics:** `str()`, vectors, data frames, factors; measurement levels in criminology context

**Learning Activities:**
- Explore crimsyndata package structures
- Practice Run/Modify tasks with guided examples
- Identify measurement levels in crime datasets

**Deliverables:**
- Dodona Set A: Data Literacy (5 exercises)
- Reflection: "What surprised you about crime data structure?"

**Tracking Metrics:**
- Item-level error rates by exercise
- Median time-on-task
- Factor vs. character confusion patterns

---

## Week 3: Descriptives & APA Basics
**Topics:** `summary()`, `table()`, grouped summaries; APA tables via flextable/apaTables

**Learning Activities:**
- Generate frequency tables for offense types
- Create grouped summaries by district/demographic variables
- Introduction to APA formatting from code

**Deliverables:**
- Dodona Set B: Descriptive Statistics (4 exercises)
- Mini-memo: 200-word interpretation of results

**Tracking Metrics:**
- % submissions passing APA checks without manual edits
- Interpretation quality rubric scores
- Time spent on formatting vs. analysis

---

## Week 4: dplyr I - Filter, Select, Pipe
**Topics:** `filter()`, `select()`, pipe operator; tidy naming and code style

**Learning Activities:**
- Filter crime data by type, location, time period
- Select relevant variables for analysis
- Chain operations using pipes

**Deliverables:**
- Dodona Set C: Data Manipulation Basics (4 exercises)
- Quiz 1 in Ufora (low-stakes, 10 questions)

**Tracking Metrics:**
- Logical operator misuse patterns
- Pipe syntax adoption rate
- Common variable selection errors

---

## Week 5: dplyr II - Grouping & Crime Rates
**Topics:** `group_by()`, `summarise()`, rate calculations per 100k; NA handling

**Learning Activities:**
- Calculate crime rates by district, adjust for population
- Handle missing data appropriately
- Create summary tables with proper denominators

**Deliverables:**
- Assignment 1: Crime Rate Analysis
  - APA-formatted table of rates by district/crime type
  - 500-word interpretation following APA style
  - Reproducible R script with comments

**Tracking Metrics:**
- APA compliance checklist completion
- Denominator error frequency
- NA handling approach accuracy

---

## Week 6: Tidy Data & Joins
**Topics:** `pivot_longer()`, `pivot_wider()`, `left_join()`, data dictionaries

**Learning Activities:**
- Reshape crime data for different analysis needs
- Join crime data with demographic/geographic information
- Create and use data dictionaries

**Deliverables:**
- Dodona Set D: Data Reshaping & Joins (3 exercises)
- swirl "Basics" module completion checkpoint

**Tracking Metrics:**
- Join key mismatch frequency
- Pivot operation success rates
- Data dictionary usage compliance

---

## Week 7: Visualization Fundamentals
**Topics:** ggplot2 grammar of graphics; bar, line, histogram, scatterplot; academic styling

**Learning Activities:**
- Create publication-quality plots
- Apply consistent labeling and styling
- Choose appropriate plot types for data

**Deliverables:**
- Dodona Set E: Visualization (3 exercises)
- Two plots with captions explaining criminological insights

**Tracking Metrics:**
- Plot readability checklist pass rate
- Appropriate plot type selection
- Caption quality and insight depth

---

## Week 8: Midterm Mini-Project
**Task:** District Crime Profile Analysis using crimsyndata

**Requirements:**
- Choose 3-4 variables from district crime data
- Produce 1 APA-formatted table + 1 publication-quality figure
- Write 1-page interpretation with criminological context
- Participate in structured peer review process

**Deliverables:**
- Complete mini-project submission
- Peer review feedback using provided rubric
- Reflection on peer feedback received

**Tracking Metrics:**
- Peer rubric agreement levels
- Common feedback themes identification
- Quality improvement from draft to final

---

## Week 9: Statistical Inference I
**Topics:** t-tests, chi-square tests; assumptions, effect sizes, interpretation guidelines

**Learning Activities:**
- Test differences in crime rates between groups
- Analyze categorical associations (offense type × demographic)
- Calculate and interpret effect sizes

**Deliverables:**
- Dodona Inferential Statistics Set (4 exercises)
- swirl "t-test & chi-square" module completion

**Tracking Metrics:**
- p-value vs. effect size interpretation accuracy
- Assumption checking compliance
- Statistical significance misinterpretation patterns

---

## Week 10: Regression I - Bivariate
**Topics:** Simple linear regression; model specification, coefficients, fit assessment, diagnostics

**Learning Activities:**
- Model crime rates as function of single predictor
- Interpret regression coefficients in context
- Assess model assumptions and fit

**Deliverables:**
- Quiz 2: Interpretation Focus (Ufora, 15 questions)
- Dodona Regression Set A (3 exercises)

**Tracking Metrics:**
- Confidence interval interpretation errors
- Coefficient interpretation accuracy
- Residual analysis completion rates

---

## Week 11: Regression II - Multiple
**Topics:** Multiple regression; confounding, multicollinearity, standardized coefficients

**Learning Activities:**
- Build multivariable models with crime data
- Check VIF values and assess multicollinearity
- Compare standardized and unstandardized coefficients

**Deliverables:**
- Dodona Set F: Multiple Regression (3 exercises)
- Regression table + narrative interpretation (APA format)

**Tracking Metrics:**
- VIF threshold understanding
- Multicollinearity identification accuracy
- Model interpretation completeness

---

## Week 12: Spatial Analysis & Reproducibility
**Topics:** sf package basics, geographic joins, choropleth maps; R Projects, here(), script organization

**Learning Activities:**
- Create simple crime maps by district
- Join crime data to geographic boundaries
- Organize reproducible analysis workflows

**Deliverables:**
- Choropleth map with brief interpretation
- Cautionary note on ecological fallacy
- Organized R project with proper file structure

**Tracking Metrics:**
- CRS/projection error frequency
- Geographic join success rates
- Project organization compliance

---

## Week 13: Final Project Launch
**Task:** Student-chosen criminology question using approved datasets

**Activities:**
- Individual consultations on project ideas
- Proposal development with guidance
- Research question refinement

**Deliverables:**
- Project proposal including:
  - Research question and rationale
  - Variables and analysis plan
  - Expected outputs (table + figure types)
  - Timeline for completion

**Tracking Metrics:**
- Proposal approval turnaround <72 hours
- FAQ compilation for common issues
- Research question feasibility assessment

---

## Week 14: Student Clinics & SPSS Bridges
**Topics:** Side-by-side SPSS↔R translations; individualized support

**Activities:**
- Drop-in clinic sessions for project support
- SPSS-to-R workflow demonstrations:
  - Select Cases → `filter()`
  - Compute Variable → `mutate()`
  - Crosstabs → `table()` + `chisq.test()`
  - Regression → `lm()` + `summary()`

**Deliverables:**
- Draft results for instructor feedback
- SPSS-to-R translation exercise completion

**Tracking Metrics:**
- Clinic attendance vs. revision quality correlation
- SPSS bridge exercise completion rates
- Common translation error patterns

---

## Week 15: Finalization Week
**Focus:** Final project completion and quality assurance

**Activities:**
- Final project submissions due Friday of Week 15
- Drop-in office hours for last-minute support
- APA formatting and reproducibility checks

**Deliverables:**
- Complete final project package:
  - Written report (4-6 pages, APA format)
  - Annotated R script with analysis code
  - APA-formatted table and publication-quality figure
  - Brief reflection on learning process

**Tracking Metrics:**
- On-time submission rate
- Reproducibility spot-check pass rate (random sample)
- APA compliance final assessment

---

## Week 16: Course Wrap-up
**Activities:**
- Optional student lightning presentations (5 min each)
- Course reflection survey and feedback collection
- Resource sharing for continued learning

**Deliverables:**
- Final course reflection survey
- Peer feedback on presentations (if participating)

**Tracking Metrics:**
- Overall course satisfaction ratings
- Self-efficacy change from baseline
- Future learning interests and confidence

---

# Success Metrics & Continuous Improvement

## Engagement Metrics
```{r engagement_metrics, echo=FALSE}
engagement <- data.frame(
  Metric = c("Onboarding Completion", "Exercise Submission Rate", 
             "Clinic Attendance", "Peer Review Participation"),
  Target = c("≥85% by Week 1 Day 5", "≥90% within deadline", 
             "≥60% when needed", "≥95% meaningful feedback"),
  Tracking = c("Daily dashboard", "Weekly reports", 
               "Session logs", "Rubric completion")
)
kable(engagement, caption = "Engagement Success Metrics")
```

## Mastery Metrics
```{r mastery_metrics, echo=FALSE}
mastery <- data.frame(
  Skill_Area = c("Data Manipulation", "Rate Calculations", "APA Formatting", 
                 "Statistical Interpretation", "Code Organization"),
  Success_Indicator = c("Filter/select/mutate accuracy", "Proper denominators + NA handling",
                        "First-pass compliance ≥70%", "Effect size vs p-value clarity",
                        "Reproducible project structure"),
  Assessment_Method = c("Dodona auto-grading", "Assignment rubrics",
                        "Automated APA checks", "Quiz + project evaluation",
                        "Code review checklist")
)
kable(mastery, caption = "Mastery Success Metrics")
```

## Quality Assurance Process
1. **Weekly Error Analysis**: Identify common mistakes → update hints/examples
2. **Student Feedback Loops**: 1-minute pulse polls at Weeks 1, 5, 8, 15
3. **Peer Review Calibration**: Sample review agreement analysis
4. **APA Compliance Tracking**: Automated checks + manual verification
5. **Language Clarity Assessment**: B1 readability + native speaker review

## Continuous Improvement Cycle
- **Bi-weekly:** Review error patterns and adjust hints
- **Mid-semester:** Comprehensive feedback analysis and content updates
- **End-of-course:** Full curriculum review and next-iteration planning
- **Post-course:** Alumni follow-up on skill retention and application

---

# Platform-Specific Implementation

## Dodona Integration
- **Auto-graded Exercises**: 15 core items + 5 bonus/extension
- **Hint System**: Multi-level hints triggered by common errors
- **Progress Tracking**: Individual and cohort-level dashboards
- **Accessibility**: Screen reader compatible, multiple input methods

## swirl Integration  
- **Custom Course**: "Bachelor Criminology Data Analysis"
- **Interactive Tutorials**: Console-based practice with immediate feedback
- **Progress Syncing**: Links to Ufora gradebook for completion tracking
- **Offline Capability**: Students can practice without internet access

## Ufora Coordination
- **Resource Hub**: Lecture slides, datasets, reference materials
- **Discussion Forums**: Peer help, Q&A, announcement space
- **Assessment Integration**: Quizzes sync with overall grade calculation
- **Communication**: Weekly announcements, deadline reminders

## GitHub Integration
- **Read-Only Access**: Students can view but not edit main repositories
- **Dataset Distribution**: Version-controlled crimsyndata releases
- **Example Scripts**: Template code for common tasks
- **Documentation**: Comprehensive guides and troubleshooting resources

---

# Risk Mitigation & Contingency Planning

## Technical Risks
- **R/RStudio Installation Issues**: Pre-recorded setup videos + virtual lab backup
- **Dodona Platform Downtime**: Offline exercise alternatives prepared
- **Package Compatibility**: Version-locked environment specifications

## Pedagogical Risks  
- **Varied Programming Background**: Multi-track approach with catch-up resources
- **Statistical Anxiety**: Low-stakes assessments + peer support structures
- **Time Management**: Clear deadlines + checkpoint system

## Language Delivery Risks
- **Dutch Comprehension**: English alternatives + translation glossary
- **Technical Terminology**: Bilingual concept mapping + definition bank
- **Cultural Context**: Local crime examples + international comparisons

## Quality Assurance Backup Plans
- **Instructor Illness**: Guest lecturer arrangements + recorded content
- **Student Support Overload**: Teaching assistant training + FAQ automation
- **Assessment Integrity**: Multiple versions + plagiarism detection protocols
